---
title: "DA final project code"
author: "Qifei Yuan"
date: "12/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preprocessing
```{r import library}
library(plyr)
library(foreign)
library(ggplot2)
library(psych)
library(ordinal)

```

```{r preprocessing}
# read Stata file

df_full <- read.dta("/Users/freda/Desktop/Final 加油鸭/DA/Data set/GSS2018.dta")

# get rid of all missing values
df <- df_full[,c('happy','class',"finrela", 'age','income16','sex','race','educ','wrkstat','health','degree','relig')]

df <- na.omit(df)
```

### Dependent Variable: encoded into numerical
* 1: not too happy
* 2 pretty happy
* 3: very happy
Don't know coded into missing value
```{r recode Y missing value}
#recode "happy": convert don't know, NA into missing values
df$happy = as.character(df$happy)
df$happy[df$happy== 'DK'] <-NA
df$happy[df$happy=='IAP'] <-NA
df$happy[df$happy=='NA'] <-NA

#convert character to numeric type
# df$happy[df$happy == 'not too happy'] <- 1
# df$happy[df$happy == 'pretty happy'] <- 2
# df$happy[df$happy == 'very happy'] <- 3
# df$happy = as.factor(df$happy)

#describe(df$happy)
```

```{r }
# happy2 = 1 if very happy
df$happy2 <- df$happy
df$happy2[df$happy=='very happy'] <- 1
df$happy2[df$happy=='pretty happy'] <- NA
df$happy2[df$happy=='not too happy'] <- 0

df$happy2 = as.factor(df$happy2)
summary(df$happy2)
```


## Recode independent variables:

### (1) ln_income: natural log of income midpoints

```{r recode income}
# recode "income": take midpoint + 1.5 times in upper limit
# get rid of missing values: start from rincom16 

df$income16 = as.character(df$income16)
df$income16[df$income16== 'dk'] <-NA
df$income16[df$income16=='IAP'] <-NA
df$income16[df$income16=='na'] <-NA
df$income16[df$income16=='refused'] <-NA

df$income16 = as.factor(df$income16)
summary(df$income16)

# create "income" (numeric), coded into midpoint
df$income = as.character(df$income16)

df$income[df$income == 'under $1 000'] <- mean(c(0,1000))
df$income[df$income == '$1 000 to 2 999'] <- mean(c(1000,2999))
df$income[df$income == '$3 000 to 3 999'] <- mean(c(3000,3999))
df$income[df$income == '$4 000 to 4 999'] <- mean(c(4000,4999))
df$income[df$income == '$5 000 to 5 999'] <- mean(c(5000,5999))
df$income[df$income == '$6 000 to 6 999'] <- mean(c(6000,6999))
df$income[df$income == '$7 000 to 7 999'] <- mean(c(7000,7999))
df$income[df$income == '$8 000 to 9 999'] <- mean(c(8000,9999))
df$income[df$income == '$10000 to 12499'] <- mean(c(10000,12499))
df$income[df$income == '$12500 to 14999'] <- mean(c(12500,14999))
df$income[df$income == '$15000 to 17499'] <- mean(c(15000,17499))
df$income[df$income == '$17500 to 19999'] <- mean(c(17500,19999))
df$income[df$income == '$20000 to 22499'] <- mean(c(20000,22499))
df$income[df$income == '$22500 to 24999'] <- mean(c(22500,24999))
df$income[df$income == '$25000 to 29999'] <- mean(c(25000,29999))
df$income[df$income == '$30000 to 34999'] <- mean(c(30000,34999))
df$income[df$income == '$35000 to 39999'] <- mean(c(35000,39999))
df$income[df$income == '$40000 to 49999'] <- mean(c(40000,49999))
df$income[df$income == '$50000 to 59999'] <- mean(c(50000,59999))
df$income[df$income == '$60000 to 74999'] <- mean(c(60000,74999))
df$income[df$income == '$75000 to $89999'] <- mean(c(75000,89999))
df$income[df$income == '$90000 to $109999'] <- mean(c(90000,109999))
df$income[df$income == '$110000 to $129999'] <- mean(c(110000,129999))
df$income[df$income == '$130000 to $149999'] <- mean(c(130000,149999))
df$income[df$income == '$150000 to $169999'] <- mean(c(150000,169999))
df$income[df$income == '$170000 or over'] <- 170000*1.5

df$income = as.numeric(df$income)
describe(df$income)

table(df$sex)

skew(df$income)
kurtosis(df$income)
```

# distribution of numeric income midpoint 
```{r}
p = ggplot(data=df, aes(df$income)) + 
  geom_histogram(bins = 40)
p
```

```{r}
# convert to natural log of income
df$ln_income = log(df$income)

p = ggplot(data=df, aes(df$ln_income)) + 
  geom_histogram(bins = 40)
p
```


```{r distribution of ln_income}
describe(df$ln_income)
```

```{r distribution of income}
describe(df$income)
```

### Controlled Variables:
### (2) Sex (female): female = 1, male = 0 
```{r recode X variables}
#recode sex as "female"
df$female = ifelse(df$sex=='female', 1, 0)
```


```{r recode married}
#recode marital status
# df$childs = ifelse(df$childs=='0', 0, 1)
# summary(df$childs)
```

```{r recode "work status"}
# recode work status
df$employed = as.character(df$wrkstat)
df$employed[df$wrkstat== 'DK'] <-NA
df$employed[df$wrkstat=='IAP'] <-NA
df$employed[df$wrkstat=='NA'] <-NA

summary(df$wrkstat)

# code working status into 3 new categories
df$employed[df$wrkstat== 'working fulltime'] <- 1
df$employed[df$wrkstat== 'working parttime'] <- 1
df$employed[df$wrkstat=='keeping house'] <- 0 
df$employed[df$wrkstat=='retired'] <- 0 
df$employed[df$wrkstat=='school'] <- 0 
df$employed[df$wrkstat=='other'] <- 0 
df$employed[df$wrkstat=='temp not working'] <- 0 
df$employed[df$wrkstat=='unempl, laid off'] <- 0 

df$employed = as.numeric(df$employed)
summary(df$employed)
```

```{r}
# df$work = as.character(df$wrkstat)
# df$work[df$wrkstat== 'DK'] <-NA
# df$work[df$wrkstat=='IAP'] <-NA
# df$work[df$wrkstat=='NA'] <-NA
# 
# # code working status into 3 new categories
# df$work[df$wrkstat== 'working fulltime'] <- "full time"
# df$work[df$wrkstat== 'working parttime'] <- "full time"
# df$work[df$wrkstat=='keeping house'] <- "other"
# df$work[df$wrkstat=='retired'] <-"other"
# df$work[df$wrkstat=='school'] <-"other"
# df$work[df$wrkstat=='other'] <-"other"
# df$work[df$wrkstat=='temp not working'] <- "temp unemployed"
# df$work[df$wrkstat=='unempl, laid off'] <- "longterm unemployed"
# 
# df$work = as.factor(df$work)
# summary(df$work)
``` 
    
```{r recode health}
df$health2 = as.character(df$health)

# code working status into 3 new categories
df$health2[df$health== 'excellent'] <- "good"
df$health2[df$health== 'good'] <- "good"
df$health2[df$health=='fair'] <- "bad"
df$health2[df$health=='poor'] <-"bad"
df$health2 = as.factor(df$health2)

summary(df$health2)


```

```{r recode finrela}
df$finrela = as.character(df$finrela)
df$finrela[df$finrela== 'DK'] <-NA
df$finrela[df$finrela=='IAP'] <-NA
df$finrela[df$finrela=='NA'] <-NA
df$finrela = as.factor(df$finrela)


df$finrela <- factor(df$finrela, levels = c('far below average','below average','average','above average','far above average'))
summary(df$finrela)
``` 


```{r recode race}
#recode race
df$white <- ifelse(df$race=='white', 1, 0)
```

```{r recode religion}
#recode religion
df$relig <- ifelse(df$relig=='none', 0, 1)
summary(df$relig)
```

## visualize relationship
```{r distribution of Y: happy}
summary(df$happy)
prop.table(table(df$happy)) 
```

### Boxplot: happy vs. ln_income
```{r boxplot}
                 
p1= ggplot(df, aes(x=factor(happy), y=ln_income, fill=factor(happy))) + 
    geom_boxplot(alpha=0.3) +
    # geom_jitter() +
    scale_fill_brewer(palette="Dark2")                 

p1                
```







```{r}
#recode happy
df$happy.num <- df$happy
df$happy.num[df$happy == 'not too happy'] <- 1
df$happy.num[df$happy == 'pretty happy'] <- 2
df$happy.num[df$happy == 'very happy'] <- 3
# df$happy.num = as.factor(df$happy.num)

#describe(df$happy)


df$happy.new <- factor(df$happy.num,
                       levels = 1:3,
                        labels = c("not too happy", "pretty happy", "very happy"),
                         ordered = TRUE)
table(df$happy.new)

table(df$happy)
```

```{r recode degree}
df$degree2 = as.character(df$degree)
df$degree2[df$degree2 == 'lt high school'] <- 'no college'
df$degree2[df$degree2 == 'high school'] <- 'no college'
df$degree2[df$degree2 == 'junior college'] <- 'college'
df$degree2[df$degree2 == 'bachelor'] <- 'college'
df$degree2[df$degree2 == 'graduate'] <- 'college'
df$degree2 = as.factor(df$degree2)

summary(df$degree2)
``` 

```{r descriptive }

describe(df)

library(pastecs)
sub <-df[,c(as.numeric("happy"),"class", "age")]
stat.desc(sub)

``` 




```{r model 1}
library(stargazer)
# linear model: should be bad
#income significant, educ not
df$agesq = df$age^2

# first, initial model:

linear1 = lm(happy.num ~ ln_income + educ + age + female + white + employed + health2 + relig, data=df) 
summary(linear1) 

linear2 = lm(happy.num ~ class + educ + age + female + white + employed + health2 + relig, data=df) 
summary(linear2) 

linear3 = lm(happy.num ~ finrela + educ + age + female + white + employed + health2 + relig, data=df) 
summary(linear3) 


stargazer(linear1,linear2,linear3, type = 'html', out = "linear reg.html", title = "Initial Linear Models")

# ordinal model 
df <- within(df, finrela <- relevel(finrela, ref = 1))

lm1 = clm(factor(happy) ~ ln_income + educ + wrkstat  + age + female + white + health2 + relig, data=df) 
summary(lm1) 

lm2 = clm(factor(happy) ~ class + educ +  age + female + white +  wrkstat + health2 + relig, data=df) 
summary(lm2)

lm3 = clm(factor(happy) ~  finrela + wrkstat + educ + age + female + white + health2 + relig, data=df) 
summary(lm3) 


stargazer(lm1,lm2,lm3, type = 'html', out = "regression 2.html", title = "test")


# intermediate steps: lots of interaction, bad
lm3 = clm(factor(happy) ~ ln_income * educ + age + female + white  + ln_income *employed + ln_income*health + relig, data=df) 
summary(lm3) 

# binary logistic: only happy vs. unhappy
lm4 = glm(factor(happy2) ~ ln_income + educ + age + female + wrkstat + health2 + white + relig, df, family=binomial)
summary(lm4)

# improved: ordinal, sig, separate work status -> now show employment 

lm3= clm(factor(happy) ~ ln_income + degree +  age + female + white + work + health + relig, data=df)
summary(lm3) 

lm_test = clm(factor(happy) ~ ln_income + degree +  age + female + white + wrkstat + health2 + relig, data=df,subset = (age > 20))
summary(lm_test) 

# good model!! 
lm_final = clm(factor(happy) ~ ln_income + degree +  age+ female + white + wrkstat + health2 + relig, data=df,subset = (age > 20 ))
summary(lm_final) 


lm_test = clm(factor(happy) ~ class +degree +  age + female + white + wrkstat + health2 + relig, data=df,subset = (age > 20))
summary(lm_test) 

```


## Correlation 
```{r check correlation}
# happy vs. income
p1= ggplot(df, aes(x=factor(happy), y = ln_income, fill=factor(happy))) + 
    geom_boxplot(alpha=0.3) +
    # geom_jitter() +
    #facet_grid(~degree) +
    scale_fill_brewer(palette="Dark2")                 

p1    


p2= ggplot(df, aes(x=finrela, y=ln_income, fill=finrela)) + 
    geom_boxplot(alpha=0.3) +
    # geom_jitter() +
    scale_fill_brewer(palette="Dark2")      
p2



p2= ggplot(df, aes(x=factor(happy), y = ln_income, fill=factor(happy))) + 
    geom_boxplot(alpha=0.3) +
    # geom_jitter() +
    #facet_grid(~degree) +
    scale_fill_brewer(palette="Dark2")                 

p2    




p3= ggplot(df, 
       aes(x = class, 
           fill = finrela)) + 
  geom_bar(position = "dodge")
p3
``` 

```{r }
# heatmap
df$happy.num <- as.numeric(df$happy.num)
summary(df$happy.num)

df.num <- df[,c('happy.num','age','ln_income','educ','white','employed','female','relig')]
matrix.data = as.matrix(df.num)
heatmap(matrix.data, scale="column")



## heatmap 2
df$class.num = as.numeric(df$class)
df$finrela.num = as.numeric(df$finrela)
df$health2.num = as.numeric(df$health2)

df.num = df[,c('happy.num','class.num','finrela.num','health2.num',"employed",'ln_income')]
library(ggcorrplot)
corr <- round(cor(df.num), 2)
corr


ggcorrplot(corr)
```